### 2017-12-11 code.Rmd
first commit

% Code

## General approach

The data com from a series of different experiments. There is one single `xlsx` file with each experiment in a single sheet. A single `csv` file was created from exporting each of those sheets. We will conduct the analysis separated for each experiment. Follows the commented code used for these analyses separated per experiment.

## Load packages

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(tidyverse)
library(readxl)
library(viridis)
library(ggthemes)
library(lme4)
library(cowplot)
library(broom)
library(agricolae)
library(lattice)
library(influence.ME)
```

## Perithecia production

### Import

Let's first import the data and have a look at the raw data.

```{r message=FALSE, warning=FALSE}
perithecia <- read_csv("./data/perithecia.csv") %>%
  filter(genotype != "t")

perithecia
```

### Transform
The data were prepared in the wide format where each score for percent perithecia coverage (0 -3) is in a separate colum. Let's transform this dataframe from the wide to the long format using `gather` function of `dplyr`. We need to create a column for the variable name and another for the score.

```{r}
perithecia2 <- perithecia %>%
  gather("score", "frequency", 8:11)

perithecia2
```

### Visualize

Let's have a look at the frequency of the scores for each species-genotype separated for each substrate. This is a good plot to show in the paper.

```{r}
g1 <- perithecia2 %>%
  ggplot(aes(reorder(genotype, -ppi), frequency, fill = score)) +
  geom_bar(stat = "identity", position = "fill") +
  facet_wrap(~substrate) +
  theme_few() + # load ggthemes
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "", y = "Proportion of grains", fill = "Score") +
  scale_fill_viridis(discrete = TRUE)
g1
```

### Model

Since we have one species which has two genotypes, it may be important, besides testing the effect of species or genotype in the model, to create a species*genotype interaction variable and fit a mixed model to test the effect of this variable on the perithecial production index (ppi), so that the two genotypes of one species can be compared.

```{r}
# creating the variable
perithecia <- perithecia %>%
  unite(species_genotype, species, genotype, sep = "_", remove = F)
```

Now we fit the model using `lmer` function. We treat `Isolate` as a random effects.

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
glmm_per <- lmer(
  ppi ~ species_genotype * substrate + (1 | Isolate),
  data = perithecia, REML = FALSE
)
```

Let's check the single and interaction effects.

```{r message=FALSE, warning=FALSE}
library(car)
Anova(glmm_per)
```

Evaluate the model

```{r}
library(lattice)
plot(glmm_per, type = c("p", "smooth"))
qqmath(glmm_per, id = 0.05)
```

#### Means comparison

Not too bad. Let's now compare the means of treatments and create a dataframe which will be use to further create a plot with the estimated means and confidence interval. We used the `emmeans` package which is an update for the old `lsmeans` package. The syntax is the same as before.

```{r message=FALSE, warning=FALSE}
library(emmeans)
medias <- emmeans(glmm_per, ~ species_genotype * substrate)
med <- cld(medias, Letters = LETTERS, alpha = .05)

# make output as a dataframe
med <- data.frame(med)
head(med)
```

### Figures

This new `med` dataframe does not contain the two original variables we want to use to plot the data. Let's separate the `species_genotypes` variable and create `genotype` and `species`.

```{r}
med2 <- med %>%
  separate(species_genotype, c("species", "genotype"), sep = "_", extra = "merge")
# extra argument was used to keep the genotype information alltogether
head(med2)
```

We now produce the plot with point estimate for the means and error bars for the 95%CI of the `ppi` for each `genotype`. We add `species` to the legend when using `color` argument.

```{r message=FALSE, warning=FALSE}
g2 <- med2 %>%
  ggplot(aes(reorder(genotype, -emmean), emmean, color = species)) +
  geom_point(position = position_dodge(width = 0.3)) +
  theme_few() +
  theme(legend.position = "right", legend.text = element_text(size = 6, face = "italic"), axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_viridis(discrete = TRUE) +
  facet_wrap(~substrate) +
  geom_errorbar(
    aes(ymin = lower.CL, ymax = upper.CL),
    width = 0.2, position = position_dodge(width = 0.3)
  ) +
  labs(y = "Perithecia production index (%)", x = "Trichothecene genotype")
g2
```

Finally, we prepare and save the Figure 1 of the manuscript which is a combo figure with the the two previous plots for both the original scores and the normalized index. We use the `plot_grid` function of the `cowplot` package for this.

```{r message=FALSE, warning=FALSE}
grid1 <- plot_grid(g1, g2, labels = c("A", "B"), align = "hv", ncol = 1)
ggsave("figs/grid1.png", grid1, width = 6, height = 6)
```

## Mycelial growth

### Import

There is one `xlsx` file with data from a single experiment. We will use `read_excel` to correctly load the centigrades symbol for the temperature factors after setting locale to `pt_BR.UTF-8`.

```{r message=FALSE, warning=FALSE}
Sys.setlocale("LC_ALL", "pt_BR.UTF-8")
mgr <- read_excel("./data/mycelia.xlsx")

head(mgr)
```

### Visualize

We will now produce the plot with the means of mycelial growth rate for each species-genotype and temperature. This will be the final plot for publication.

```{r message=FALSE, warning=FALSE}
mgr1 <- mgr %>%
  group_by(isolate, genotype, species, temperature) %>%
  summarize(mean_mgr = mean(mgr)) %>%
  ggplot(aes(factor(genotype), mean_mgr, color = species)) +
  geom_jitter(width = 0.1, size = 3) +
  facet_wrap(~temperature) +
  scale_color_viridis(discrete = TRUE) +
  theme_few() +
  theme(
    legend.position = "none",
    legend.text = element_text(size = 7, face = "italic")
  ) +
  labs(
    x = "Trichothecene Genotype",
    y = "Mycelial growth rate (mm/day)", color = "Species"
  ) +
  ylim(0.2, 1.8)
```

### Model

Let's fit a mixed model for the `mgr` data. We first test the effect of the interaction.

```{r}
mgr <- mgr %>%
  unite(species_genotype, species, genotype, sep = "_", remove = F)
lmer_mgr <- lmer(mgr ~ species_genotype * temperature + (1 | species_genotype / isolate), data = mgr, REML = FALSE)
Anova(lmer_mgr)
```

We now create different datasets for each temperature and test the effect of `species_genotype` within each temperature.

```{r}
# 15 C
mgr15 <- mgr %>%
  filter(temperature == "15 °C")

lmer_mgr15 <- lmer(mgr ~ species_genotype + (1 | species_genotype / isolate), data = mgr15, REML = FALSE)

Anova(lmer_mgr15)

# 30 C
mgr30 <- mgr %>%
  filter(temperature == "30 °C")

lmer_mgr30 <- lmer(mgr ~ species_genotype + (1 | species_genotype / isolate), data = mgr30, REML = FALSE)

Anova(lmer_mgr30)
```

We can see that there are no effect of `species_genotype` in the `mgr` within each of the temperatures.

## Sporoulation & Germination

In this experiment, the production of macroconidia and further germination of 20 randomly selected spores of the different strains of the species-genotype combination, were evaluated in two runs of the experiment.

### Import

```{r message=FALSE, warning=FALSE}
spor <- read_csv("./data/spor.csv")

# again, we create the species_genotype variable
spor <- spor %>%
  unite(species_genotype, species, genotype, sep = "_", remove = F)
head(spor)
```

### Visualize

The code below will produce plots for total macroconidia production and percent germinated spores.

```{r message=FALSE, warning=FALSE}
library(scales)
spor1 <- spor %>%
  group_by(species, genotype, isolate) %>%
  summarize(mean_spor = mean(spores_ml))
spor2 <- spor1 %>%
  ggplot(aes(genotype, mean_spor, color = species)) +
  geom_jitter(width = 0.1, size = 3) +
  theme_few() +
  scale_color_viridis(discrete = TRUE) +
  theme(legend.position = c(1, 5), axis.text.x = element_text(angle = 0, hjust = 0.5, size = 7), legend.text = element_text(size = 7, face = "italic"), axis.title = element_text(size = 8)) +
  labs(x = "", y = "Macroconidia count (x 10.000)", color = "Species") +
  ylim(0, 5) +
  guides(colour = guide_legend(nrow = 2))
```

```{r message=FALSE, warning=FALSE}
germ1 <- spor %>%
  group_by(species, genotype, isolate) %>%
  summarize(mean_germ = mean(germp))
germ <- germ1 %>%
  ggplot(aes(genotype, mean_germ, color = species)) +
  geom_jitter(width = 0.1, size = 3) +
  theme_few() +
  scale_color_viridis(discrete = TRUE) +
  theme(legend.position = "none", axis.text.x = element_text(angle = 0, hjust = 0.5, size = 7), legend.text = element_text(size = 7, face = "italic"), axis.title = element_text(size = 8)) +
  labs(x = "", y = "Germination rate (%)", color = "Species") +
  guides(colour = guide_legend(nrow = 2))
```

### Model

Anova model

```{r message=FALSE, warning=FALSE}
lmer_spor <- lmer(spores_ml ~ species_genotype + (1 | species_genotype), data = spor, REML = FALSE)
Anova(lmer_spor)
plot(lmer_spor, type = c("p", "smooth"))
qqmath(lmer_spor, id = 0.05)
```

#### Means comparison

```{r}
medias <- emmeans(lmer_spor, ~species_genotype)
med <- cld(medias, Letters = LETTERS, alpha = .05)
med <- data.frame(med)
med
```

### Figure

```{r message=FALSE, warning=FALSE}
grid1 <- plot_grid(spor2, germ, labels = c("B", "C"), ncol = 2, align = "hv")

grid5 <- plot_grid(mgr1, grid1, labels = c("A"), rel_heights = c(1, 1), ncol = 1, align = "hv")
ggsave("figs/mgr_spor_germ.png", grid5, width = 5, height = 5)

grid5
```

## Pathogenicity tests

### Import

```{r message=FALSE, warning=FALSE}
pathogen <- read_csv("./data/pathogenicity.csv")
pathogen <- pathogen %>%
  filter(isolate != "test")
```

### Transform

Let's prepare several summary statistics for the disease variables per cultivar. We will calculate the mean and standard deviation of proportion of diseased spikelets.

```{r}
# preparing cultivar 194

pathogen_194 <- pathogen %>%
  filter(cultivar == "BRS 194") %>%
  select(1:7, 9, 11, 12, 13) %>%
  gather(position, count, 8:9) %>%
  # filter(exp == 2) %>%
  group_by(cultivar, dai, genotype, species) %>%
  summarize(
    mean_sev = mean(sev),
    sd_sev = sd(sev)
  )

# preparing Guamirim

pathogen_Gua <- pathogen %>%
  filter(cultivar == "BRS Guamirim") %>%
  select(1:7, 9, 11:13) %>%
  gather(position, count, 8:9) %>%
  # filter(exp == 2) %>%
  group_by(cultivar, dai, genotype, species, position) %>%
  summarize(
    mean_sev = mean(sev),
    sd_sev = sd(sev)
  )
```

Let's calculate the area under the curve for each isolate. For this, we will use the `audpc` function of the `agricolae` package. First, we need to group several variables up to `dai` (days after inoculation). To make our life easier, lets use the `do` in conjuction with the `tidy` functions of the `dplyr` and `broom` package, respectively. This will calculate the audpce for each spike.

```{r}
audpc_194 <- pathogen %>%
  filter(cultivar == "BRS 194") %>%
  # filter(exp == 2) %>%
  unite(species_genotype, species, genotype, sep = "_", remove = F) %>% 
  select(
    exp, cultivar, dai, isolate, species_genotype, species, genotype, spike,
    sev
  ) %>%
  group_by(exp, isolate, species_genotype, species, genotype, spike, dai) %>%
  summarize(mean_sev = mean(sev)) %>%
  do(tidy(audpc(.$mean_sev, .$dai)))
names(audpc_194)[8] <- "audpc"
```

```{r}
audpc_gua <- pathogen %>%
  filter(cultivar == "BRS Guamirim") %>%
  # filter(exp == 2) %>%
   unite(species_genotype, species, genotype, sep = "_", remove = F) %>% 
  select(
    exp, cultivar, dai, isolate, species_genotype, species, genotype, spike,
    sev
  ) %>%
  group_by(exp, isolate, species_genotype, species, genotype, spike, dai) %>%
  summarize(mean_sev = mean(sev)) %>%
  do(tidy(audpc(.$mean_sev, .$dai)))
names(audpc_gua)[8] <- "audpc"
```

### Visualize

```{r}

audpc_194 %>%
  group_by(species, genotype) %>%
  summarize(
    mean_audpc = mean(audpc),
    sd_audpc = sd(audpc)
  ) %>%
  ggplot(aes(species, mean_audpc, color = genotype)) +
  coord_flip() +
  theme_few() +
  geom_point(position = position_dodge(width = 0.4)) +
  geom_errorbar(aes(min = mean_audpc - sd_audpc, max = mean_audpc + sd_audpc), position = position_dodge(width = 0.4), width = 0.2)
```

```{r}
audpc_gua %>%
  group_by(species, genotype) %>%
  summarize(
    mean_audpc = mean(audpc),
    sd_audpc = sd(audpc)
  ) %>%
  ggplot(aes(species, mean_audpc, color = genotype)) +
  coord_flip() +
  theme_few() +
  geom_point(position = position_dodge(width = 0.4)) +
  geom_errorbar(aes(min = mean_audpc - sd_audpc, max = mean_audpc + sd_audpc), position = position_dodge(width = 0.4), width = 0.2)
```

### Model

#### BRS Guamirim

```{r}

mix_gua <- lmer(
  audpc ~ genotype +
    (1 | exp/spike),
  data = audpc_gua, REML = FALSE
)
```

Evaluate the model

```{r message=FALSE, warning=FALSE}

plot(mix_gua, type = c("p", "smooth"))
qqmath(mix_gua, id = 0.05)
```

```{r}
Anova(mix_gua)

library(emmeans)
means <- emmeans(mix_gua, ~ genotype)
cld(means)
```

```{r}
mix_194 <- lmer(log(audpc) ~ genotype + (1 | exp / spike), data = audpc_194, REML = FALSE)
```

Evaluate the model

```{r}

plot(mix_194, type = c("p", "smooth"))
qqmath(mix_194, id = 0.05)
```

#### BRS 194

```{r}
Anova(mix_194)
library(emmeans)
means <- emmeans(mix_194, ~ genotype)
cld(means)
```