### 2017-12-11 index.Rmd

% About

<img src = "https://www.apsnet.org/edcenter/intropp/lessons/fungi/ascomycetes/PublishingImages/Fusariumdiscyclesm.jpg" align= right width = 250> This is a fully reproducible report that contains and describe the data and the R codes for analyzing and visualizing these data obtained from a series of experiments conducted as an MSc research in the Del Ponte lab. The research's objectives were to assess and compare members of a species complex (FGSC), with varying toxigenic potential, that cause [Fusarium head blight or scab](https://www.apsnet.org/edcenter/intropp/lessons/fungi/ascomycetes/Pages/Fusarium.aspx), an important disease of wheat and barley worldwide. Several fitness-related traits were compared in five experiments conducted both *in vitro* and in *planta*. Data from each experiment were analyzed separately.

## Summary of methods

A collection of 30 strains was obtained from the laboratory's FGSC collection and experiments were designed to measure traits related with asexual and sexual fungal growth, pathogenicity towards two wheat cultivars of varying susceptibility and sensitivity to tebuconazole fungicide. In most analyses, we were interested in differences among the species-genotype and not on individual isolates. Therefore, the isolates were treated as random sample representative of the population of strains. In the models, they were treated as random effects (mixed-modeling).

## Project team

Project PI:

- Emerson M. Del Ponte
- Universidade Federal de Viçosa

Members:

- Camila P. Nicolli, MSc student
- Piérri Spolti, Postdoctoral felow
- Franklin Machado, DSc student

## Citation

Please cite this research compendium as:

> Del Ponte EM. Title of compendium. Online at https://doi.org/xxx/xxx

## Licences

Data: [CC-0](https://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse
Manuscript: [CC-BY-4.0](https://creativecommons.org/licenses/by/4.0/)  
Code: [MIT](https://opensource.org/licenses/MIT) year: 2017, copyright holder: Emerson Del Ponte

## Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.html). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.html). By participating in this project you agree to abide by its terms.