### 2017-12-16 code.Rmd
new changes

% Code

## Introduction

The data com from a series of different experiments. There is one single `xlsx` file with each experiment in a single sheet. Follows the commented code used for these analyses separated per experiment.

## Load packages

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(tidyverse)
library(here)
library(viridis)
library(ggthemes)
library(lme4)
library(cowplot)
library(broom)
library(agricolae)
library(lattice)
library(emmeans)
library(car)
library(scales)
```

## Perithecia production

### Import

Let's first import the data and have a look at the raw data.

```{r message=FALSE, warning=FALSE}
perithecia <- read_csv(here("data", "perithecia.csv")) %>%
  filter(genotype != "t")

perithecia
```

### Transform

The data were prepared in the wide format where each score for percent perithecia coverage (0 -3) is in a separate column. Let's transform this dataframe from the wide to the long format using `gather` function of `dplyr`. We need to create a column for the variable name and another for the score.

```{r}
perithecia2 <- perithecia %>%
  gather("score", "frequency", 8:11)

perithecia2
```

### Visualize

Let's visualize the frequency of the scores for each species-genotype separated for each substrate. This is a nice plot to show in the paper!

```{r}
g1 <- perithecia2 %>%
  ggplot(aes(reorder(genotype, -ppi), frequency, fill = score)) +
  geom_bar(stat = "identity", position = "fill") +
  facet_wrap(~substrate) +
  theme_few() + # load ggthemes
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "", y = "Proportion of grains", fill = "Score") +
  scale_fill_viridis(discrete = TRUE)
g1
```

### Model

Since we have one species which has two genotypes, it may be important, besides testing the effect of species or genotype in the model, to create a species*genotype interaction variable and fit a mixed model to test the effect of this variable on the perithecial production index (ppi). This way the two genotypes within one species can be compared between them and with the others.

```{r}
# creating the variable
perithecia <- perithecia %>%
  unite(species_genotype, species, genotype, sep = "_", remove = F)
```

Now we fit the model using `lmer` function. We treat `Isolate` as a random effects.

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
glmm_per <- lmer(
  ppi ~ species_genotype * substrate + (1 | Isolate),
  data = perithecia, REML = FALSE
)
```

Let's check the single and interaction effects.

```{r message=FALSE, warning=FALSE}
Anova(glmm_per)
```

Evaluate the model

```{r}
plot(glmm_per, type = c("p", "smooth"))
qqmath(glmm_per, id = 0.05)
```

#### Means comparison

Not too bad. Let's now compare the means of treatments and create a dataframe which will be use to further create a plot with the estimated means and confidence interval. We used the `emmeans` package which is an update for the old `lsmeans` package. The syntax is the same as before.

```{r message=FALSE, warning=FALSE}
medias <- emmeans(glmm_per, ~ species_genotype * substrate)
med <- cld(medias, Letters = LETTERS, alpha = .05)

# make output as a dataframe
med <- data.frame(med)
head(med)
```

### Figures

This new `med` dataframe does not contain the two original variables we want to use to plot the data. Let's separate the `species_genotype` variable and create `genotype` and `species`.

```{r}
med2 <- med %>%
  separate(species_genotype, c("species", "genotype"), sep = "_", extra = "merge")
# extra argument was used to keep the genotype information alltogether
head(med2)
```

We now produce the plot with point estimate for the means and error bars for the 95%CI of the `ppi` for each `genotype`. We add `species` to the legend when using `color` argument.

```{r message=FALSE, warning=FALSE}
g2 <- med2 %>%
  ggplot(aes(reorder(genotype, -emmean), emmean, color = species)) +
  geom_point(position = position_dodge(width = 0.3)) +
  theme_few() +
  theme(legend.position = "right", legend.text = element_text(size = 8, face = "italic"), axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_viridis(discrete = TRUE) +
  facet_wrap(~substrate) +
  geom_errorbar(
    aes(ymin = lower.CL, ymax = upper.CL),
    width = 0.2, position = position_dodge(width = 0.3)
  ) +
  labs(y = "Perithecia production index (%)", x = "Trichothecene genotype")
g2
```

Finally, we prepare and save the Figure 1 of the manuscript which is a combo figure with the the two previous plots for both the original scores and the normalized index. We use the `plot_grid` function of the `cowplot` package for this.

```{r message=FALSE, warning=FALSE}
grid1 <- plot_grid(g1, g2, labels = c("A", "B"), align = "hv", ncol = 1)
ggsave(here("figs", "figure1.png"), grid1, width = 6, height = 7)
grid1
```

## Mycelial growth

### Import

There is one `csv` file with data from a single experiment. We will use `read_csv` to impor the data.

```{r message=FALSE, warning=FALSE}

mgr <- read_csv(here("data", "mycelia.csv"))

head(mgr)
```

### Visualize

We will now produce the plot with the means of mycelial growth rate for each species-genotype and temperature. This will be the final plot for publication.

```{r message=FALSE, warning=FALSE}
mgr1 <- mgr %>%
  group_by(isolate, genotype, species, temperature) %>%
  summarize(mean_mgr = mean(mgr)) %>%
  ggplot(aes(factor(genotype), mean_mgr, color = species)) +
  geom_jitter(width = 0.1, size = 3) +
  facet_wrap(~temperature) +
  scale_color_viridis(discrete = TRUE) +
  theme_few() +
  theme(
    legend.position = "none",
    legend.text = element_text(size = 7, face = "italic")
  ) +
  labs(
    x = "Trichothecene Genotype",
    y = "MGR (mm/day)", color = "Species"
  ) +
  ylim(0.2, 1.8)
```

### Model

Let's fit a mixed model for the `mgr` data. We first test the effect of the interaction.

```{r}
mgr <- mgr %>%
  unite(species_genotype, species, genotype, sep = "_", remove = F)
lmer_mgr <- lmer(mgr ~ species_genotype * temperature + (1 | species_genotype / isolate), data = mgr, REML = FALSE)
Anova(lmer_mgr)
```

We now create different datasets for each temperature and test the effect of `species_genotype` within each temperature.

```{r}
# 15 C
mgr15 <- mgr %>%
  filter(temperature == "15")

lmer_mgr15 <- lmer(mgr ~ species_genotype + (1 | species_genotype / isolate), data = mgr15, REML = FALSE)

Anova(lmer_mgr15)

# 30 C
mgr30 <- mgr %>%
  filter(temperature == "30")

lmer_mgr30 <- lmer(mgr ~ species_genotype + (1 | species_genotype / isolate), data = mgr30, REML = FALSE)

Anova(lmer_mgr30)
```

We can see that there are no effect of `species_genotype` in the `mgr` within each of the temperatures.

## Sporoulation & Germination

In this experiment, the production of macroconidia and further germination of 20 randomly selected spores of the different strains of the species-genotype combination, were evaluated in two runs of the experiment.

### Import

```{r message=FALSE, warning=FALSE}
spor <- read_csv(here("data", "spor_germ.csv"))

# again, we create the species_genotype variable
spor <- spor %>%
  unite(species_genotype, species, genotype, sep = "_", remove = F)
head(spor)
```

### Visualize

The code below will produce plots for total macroconidia production and percent germinated spores.

```{r message=FALSE, warning=FALSE}
spor1 <- spor %>%
  group_by(species, genotype, isolate) %>%
  summarize(mean_spor = mean(spores_ml))
spor2 <- spor1 %>%
  ggplot(aes(genotype, mean_spor, color = species)) +
  geom_jitter(width = 0.1, size = 3) +
  theme_few() +
  scale_color_viridis(discrete = TRUE) +
  theme(legend.position = c(1, 5), axis.text.x = element_text(angle = 0, hjust = 0.5, size = 7), legend.text = element_text(size = 12, face = "italic"), axis.title = element_text(size = 12)) +
  labs(x = "", y = "N. of macroconidia (x10.000)", color = "Species") +
  ylim(0, 5) +
  guides(colour = guide_legend(nrow = 2))
```

```{r message=FALSE, warning=FALSE}
germ1 <- spor %>%
  group_by(species, genotype, isolate) %>%
  summarize(mean_germ = mean(germp))
germ <- germ1 %>%
  ggplot(aes(genotype, mean_germ, color = species)) +
  geom_jitter(width = 0.1, size = 3) +
  theme_few() +
  scale_color_viridis(discrete = TRUE) +
  theme(legend.position = "none", axis.text.x = element_text(angle = 0, hjust = 0.5, size = 7), legend.text = element_text(size = 12, face = "italic"), axis.title = element_text(size = 12)) +
  labs(x = "", y = "Germination rate (%)", color = "Species") +
  guides(colour = guide_legend(nrow = 2))
```

### Model

Anova model

```{r message=FALSE, warning=FALSE}
lmer_spor <- lmer(spores_ml ~ species_genotype + (1 | species_genotype), data = spor, REML = FALSE)
Anova(lmer_spor)
plot(lmer_spor, type = c("p", "smooth"))
qqmath(lmer_spor, id = 0.05)
```

#### Means comparison

```{r}
medias <- emmeans(lmer_spor, ~species_genotype)
med <- cld(medias, Letters = LETTERS, alpha = .05)
med <- data.frame(med)
med
```

### Figure 2

```{r message=FALSE, warning=FALSE}
grid1 <- plot_grid(spor2, germ, labels = c("B", "C"), ncol = 2, align = "hv")

grid5 <- plot_grid(mgr1, grid1, labels = c("A"), rel_heights = c(1, 1), ncol = 1, align = "hv")
ggsave(here("figs", "figure2.png"), grid5, width = 6, height = 6.2)

grid5
```

## Pathogenicity tests

### Import

```{r message=FALSE, warning=FALSE}
pathogen <- read_csv(here("data", "pathogenicity.csv")) %>%
  filter(isolate != "test")
```

### Transform

Let's prepare several summary statistics for the disease variables per cultivar. We will calculate the mean and standard deviation of proportion of diseased spikelets.

```{r}
# preparing cultivar 194

pathogen_194 <- pathogen %>%
  filter(cultivar == "BRS 194") %>%
  select(1:7, 9, 11, 12, 13) %>%
  gather(position, count, 8:9) %>%
  # filter(exp == 2) %>%
  group_by(cultivar, dai, genotype, species) %>%
  summarize(
    mean_sev = mean(sev),
    sd_sev = sd(sev)
  )

# preparing Guamirim

pathogen_Gua <- pathogen %>%
  filter(cultivar == "BRS Guamirim") %>%
  select(1:7, 9, 11:13) %>%
  gather(position, count, 8:9) %>%
  # filter(exp == 2) %>%
  group_by(cultivar, dai, genotype, species, position) %>%
  summarize(
    mean_sev = mean(sev),
    sd_sev = sd(sev)
  )
```

### Visualize temporal

```{r message=FALSE, warning=FALSE}

brs194 <- pathogen_194 %>%
  group_by(dai, genotype, species) %>%
  ggplot(aes(dai, mean_sev, color = species, shape = genotype, group = interaction(species, genotype))) +
  geom_point(position = position_dodge(width = 0.9)) +
  geom_line() +
  geom_errorbar(
    aes(min = mean_sev - sd_sev, max = mean_sev + sd_sev),
    width = 0.2, alpha = 0.3, position = position_dodge(width = 0.9)
  ) +
  theme_few() +
  ylim(0, 55) +
  xlim(0, 25) +
  theme(legend.position = "right", legend.text = element_text(size = 9, face = "italic")) +
  scale_color_viridis(discrete = TRUE) +
  labs(
    y = "FHB severity (%) ", x = "Days after inoculation",
    color = "species"
  )

brsgua <- pathogen_Gua %>%
  group_by(dai, genotype, species) %>%
  ggplot(aes(dai, mean_sev, color = species, shape = genotype, group = interaction(species, genotype))) +
  geom_point(position = position_dodge(width = 0.9)) +
  geom_line() +
  geom_errorbar(
    aes(min = mean_sev - sd_sev, max = mean_sev + sd_sev),
    width = 0.2, alpha = 0.3, position = position_dodge(width = 0.9)
  ) +
  theme_few() +
  theme(legend.position = "none") +
  scale_color_viridis(discrete = TRUE) +
  ylim(0, 55) +
  xlim(0, 25) +
  labs(
    y = "FHB severity (%) ", x = "Days after inoculation",
    color = "species"
  )

fig3 <- plot_grid(brsgua, brs194, labels = c("A", "B"), rel_widths = c(0.7, 1), ncol = 2, align = "hv")
ggsave(here("figs", "figure3.png"), fig3, width = 9, height = 4)

fig3
```

### Visualize AUDPC

Let's calculate the area under the curve for each isolate. For this, we will use the `audpc` function of the `agricolae` package. First, we need to group several variables up to `dai` (days after inoculation). To make our life easier, lets use the `do` in conjuction with the `tidy` functions of the `dplyr` and `broom` package, respectively. This will calculate the audpce for each spike.

```{r}
audpc_194 <- pathogen %>%
  filter(cultivar == "BRS 194") %>%
  # filter(exp == 2) %>%
  unite(species_genotype, species, genotype, sep = "_", remove = F) %>%
  select(
    exp, cultivar, dai, isolate, species_genotype, species, genotype, spike,
    sev
  ) %>%
  group_by(exp, isolate, species_genotype, species, genotype, spike, dai) %>%
  summarize(mean_sev = mean(sev)) %>%
  do(tidy(audpc(.$mean_sev, .$dai)))
names(audpc_194)[8] <- "audpc"
```

```{r}
audpc_gua <- pathogen %>%
  filter(cultivar == "BRS Guamirim") %>%
  # filter(exp == 2) %>%
  unite(species_genotype, species, genotype, sep = "_", remove = F) %>%
  select(
    exp, cultivar, dai, isolate, species_genotype, species, genotype, spike,
    sev
  ) %>%
  group_by(exp, isolate, species_genotype, species, genotype, spike, dai) %>%
  summarize(mean_sev = mean(sev)) %>%
  do(tidy(audpc(.$mean_sev, .$dai)))
names(audpc_gua)[8] <- "audpc"
```

```{r}
audpc_194 %>%
  group_by(species, genotype) %>%
  summarize(
    mean_audpc = mean(audpc),
    sd_audpc = sd(audpc)
  ) %>%
  ggplot(aes(species, mean_audpc, color = genotype)) +
  coord_flip() +
  theme_few() +
  geom_point(position = position_dodge(width = 0.4)) +
  geom_errorbar(aes(min = mean_audpc - sd_audpc, max = mean_audpc + sd_audpc), position = position_dodge(width = 0.4), width = 0.2)
```

```{r}
audpc_gua %>%
  group_by(species, genotype) %>%
  summarize(
    mean_audpc = mean(audpc),
    sd_audpc = sd(audpc)
  ) %>%
  ggplot(aes(species, mean_audpc, color = genotype)) +
  coord_flip() +
  theme_few() +
  geom_point(position = position_dodge(width = 0.4)) +
  geom_errorbar(aes(min = mean_audpc - sd_audpc, max = mean_audpc + sd_audpc), position = position_dodge(width = 0.4), width = 0.2)
```

### Model

#### BRS Guamirim

```{r}

mix_gua <- lmer(
  audpc ~ genotype +
    (1 | exp / spike),
  data = audpc_gua, REML = FALSE
)
```

Evaluate the model

```{r message=FALSE, warning=FALSE}

plot(mix_gua, type = c("p", "smooth"), col = "black")
qqmath(mix_gua, id = 0.05, col = "black")
```

```{r}
Anova(mix_gua)

means <- emmeans(mix_gua, ~ genotype)
cld(means)
```

```{r}
mix_194 <- lmer(log(audpc) ~ genotype + (1 | exp / spike), data = audpc_194, REML = FALSE)
```

Evaluate the model

```{r}

plot(mix_194, type = c("p", "smooth"), col = "black")
qqmath(mix_194, id = 0.05, col = "black")
```

#### BRS 194

```{r}
Anova(mix_194)
means <- emmeans(mix_194, ~ genotype)
cld(means)
```

## Fungicide sensitivity

### Import

```{r message=FALSE, warning=FALSE}
ec50 <- read_csv(here("data", "ec50.csv"))
```

### Summarize

```{r}
ec502 <- ec50 %>%
  group_by(isolate, species, genotype) %>%
  summarize(
    mean_ec50 = mean(ec50),
    sd_ec50 = sd(ec50)
  )
```

### Visualize

Let's have a look at the mean EC50 for the isolates

```{r}
ec502 %>%
  ggplot(aes(genotype, mean_ec50, color = species)) +
  geom_jitter(width = 0.1) +
  # geom_errorbar(aes(min = mean_ec50 - sd_ec50, max = mean_ec50 + sd_ec50), width=0.2, position = position_dodge(width = 0.2))+
  theme_few() +
  theme(legend.position = "right", legend.text = element_text(size = 7, face = "italic")) +
  ylim(0, 1.8) +
  scale_color_viridis(discrete = TRUE) +
  labs(y = (expression(paste("EC"[50], " (", mu, "g/ml)"))), x = "Trichothecene genotype", color = "Species") +
  ggsave(here("figs", "ec50-new.png"), width = 4.5, height = 3)
```