### 2017-12-19 data.Rmd
Added OSF link

% Data

## Raw data

The data were obtained from five different (mostly repeated in time) experiments, and were organized in a single `xlsx` file. Data from each experiment are in a separate sheet and were exported to a `csv` file.

The data are already organized in a rectangular (tidy) format, mostly in the long-format (a raw is an observation), but some are in the wide (observations in multiple colums), and so will need some wrangling. This munging is done within the respective section of the data analysis page, after data import.

## Download

The data are stored in the `data` folder of the GitHub repository.
https://github.com/phytopathology/paper-FGSC-fitness/tree/master/data

It is also publicly availabe at the Open Science Framerwork  
Identifiers: DOI 10.17605/OSF.IO/C2MBR | ARK c7605/osf.io/c2mbr  
https://osf.io/c2mbr/