### 2017-12-20 data.Rmd
Small changes

% Data

## Raw data

The data for this study were obtained from five different (mostly repeated in time) experiments, and were organized in a single `xlsx` file. Data from each experiment are in a separate sheet and were exported to a `csv` file. They are already organized in a rectangular (tidy) format, mostly in the long-format (a raw is an observation). For some experiments, they are in the wide format (observations in multiple colums), and so will need some wrangling. This munging is done within the respective section of the data analysis page, after data import.

## Download

> The data are stored in the following repositories:

https://github.com/phytopathology/paper-FGSC-fitness/tree/master/dat  
https://osf.io/c2mbr/ 
Identifier: DOI 10.17605/OSF.IO/C2MBR