### 2018-01-21 code.Rmd
Check spell and text edit

% Analysis code

## Introduction

The data were obtained from a series of randomized replicated experiments. There is one single `csv` file that contains data from each experiment. All analyses are presented below and separated for each experiment. I follow the steps (not in that exactly order) illustrated in the R for Data Science book (http://r4ds.had.co.nz/) (Grolemund and Wickham 2017). They are: data import and tidy, explore (transform, visualize and model) and communicate.

These are the packages used.

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(tidyverse)
library(here)
library(viridis)
library(ggthemes)
library(lme4)
library(cowplot)
library(broom)
library(agricolae)
library(lattice)
library(emmeans)
library(car)
library(scales)
```

## Perithecia production

### Import

Let's have a look at the raw data.

```{r message=FALSE, warning=FALSE}
perithecia <- read_csv(here("data", "perithecia.csv")) %>%
  filter(genotype != "t")

perithecia
```

The main response variable is ppi (perithecia production index), which was already calculated. It is an index based on the frequency of the scores given by given by: PPI = ((n0× 0)+(n1 × 1)+(n2 × 2)+(n3 × 3)) × 100) / (ntotal × 3) where nx = number of grains in the respective x category (x = 1 to 4 score) and ntotal = total number of grains (25 in this study). However, the original scores are included and consist of the number of grains on each score (0 to 3).

### Transform

The frequency of each score is shown in separate columns. To plot and further work with the data, we will need to reshape the data from the wide to the long format, where each frequency of the scores need to be in a single column. We will use the `gather` function of `dplyr`. For this, we need to inform the name of columns for the `score` and `frequency` variables.

```{r}
perithecia2 <- perithecia %>%
  gather("score", "frequency", 8:11)

perithecia2
```

### Visualize

We can now visualize the frequency of the scores for each species-genotype separated (facet) for each substrate. This is a nice plot to have in the paper!

```{r}
g1 <- perithecia2 %>%
  ggplot(aes(reorder(genotype, -ppi), frequency, fill = score)) +
  geom_bar(stat = "identity", position = "fill") +
  facet_wrap(~substrate) +
  theme_few() + # load ggthemes
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "", y = "Proportion of grains", fill = "Score") +
  scale_fill_viridis(discrete = TRUE)
g1
```

### Model

Since we have one species with two genotypes, it may be important, besides testing the effect of species or genotype in the model, to create a `species_genotype` variable. We will then fit a mixed model to test the effect species_genotype on the perithecial production index (ppi). In this way the two genotypes within one species can be compared between them and among the others. We use the `unite` function to create this variable.

```{r}
# creating the variable
perithecia <- perithecia %>%
  unite(species_genotype, species, genotype, sep = "_", remove = F)
```

Now we can fit the mixed model using `lmer` function. We treat `Isolate` as a random effects.

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(lme4)
lmm_per <- lmer(
  ppi ~ species_genotype * substrate + (1 | Isolate),
  data = perithecia, REML = FALSE
)
```

Let's check the single and interaction effects using `Anova` function of the `car` package.

```{r message=FALSE, warning=FALSE}
library(car)
Anova(lmm_per)
```

Evaluate the model

```{r}
plot(lmm_per, type = c("p", "smooth"), col = "black")
qqmath(lmm_per, id = 0.05, col = "black")
```

#### Means comparison

Not too bad. Let's now compare the means of treatments and create a data-frame which will be use to further create a plot with the estimated means and confidence interval. We will use the `emmeans` package which is an update for the old `lsmeans` package. The syntax is the same as before.

```{r message=FALSE, warning=FALSE}
library(emmeans)
medias <- emmeans(lmm_per, ~ species_genotype * substrate)
med <- cld(medias, Letters = LETTERS, alpha = .05)

# make output as a dataframe
med <- data.frame(med)
head(med)
```

### Figures

This new `med` data-frame does not contain the two original variables we want to use to plot the data. Let's separate the `species_genotype` variable and create `genotype` and `species`. We use the `separate` function.

```{r}
med2 <- med %>%
  separate(species_genotype, c("species", "genotype"), sep = "_", extra = "merge")
# extra argument was used to keep the genotype information alltogether
head(med2)
```

We now can plot the point estimates for the means and respective 95%CI of the `ppi` for each `genotype`. We add `species` legends when using `color` argument so we can have five species_genotype.

```{r message=FALSE, warning=FALSE}
g2 <- med2 %>%
  ggplot(aes(reorder(genotype, -emmean), emmean, color = species)) +
  geom_point(position = position_dodge(width = 0.3)) +
  theme_few() +
  theme(legend.position = "right", legend.text = element_text(size = 8, face = "italic"), axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_viridis(discrete = TRUE) +
  facet_wrap(~substrate) +
  geom_errorbar(
    aes(ymin = lower.CL, ymax = upper.CL),
    width = 0.2, position = position_dodge(width = 0.3)
  ) +
  labs(y = "Perithecia production index (%)", x = "Trichothecene genotype")
g2
```

Finally, we prepare and save the Figure 1 of the manuscript which is a combo figure with the the two previous plots for both the original scores and the normalized index. We use the `plot_grid` function of the `cowplot` package for this.

```{r message=FALSE, warning=FALSE}
grid1 <- plot_grid(g1, g2, labels = c("A", "B"), align = "hv", ncol = 1)
ggsave(here("figs", "figure1.png"), grid1, dpi = 600, width = 6, height = 7)
grid1
```

## Mycelial growth

### Import

The main variable is mycelia growth rate, or the ratio of the growth at the fifth day and number of days. Besides the species and genotype factors, there are two levels of temperatures being tested.

```{r message=FALSE, warning=FALSE}

mgr <- read_csv(here("data", "mycelia.csv"))

head(mgr)
```

### Visualize

We can plot means of mycelial growth rate for each species-genotype and temperature as faceting variable. This plot will be included in the publication.

```{r message=FALSE, warning=FALSE}
mgr1 <- mgr %>%
  group_by(isolate, genotype, species, temperature) %>%
  summarize(mean_mgr = mean(mgr)) %>%
  ggplot(aes(factor(genotype), mean_mgr, color = species)) +
  geom_jitter(width = 0.1, size = 3) +
  facet_wrap(~temperature) +
  scale_color_viridis(discrete = TRUE) +
  theme_few() +
  theme(
    legend.position = "top",
    legend.text = element_text(size = 7, face = "italic")
  ) +
  labs(
    x = "Trichothecene Genotype",
    y = "MGR (mm/day)", color = "Species"
  ) +
  ylim(0.2, 1.8)
```

### Model

Let's fit a mixed model for the `mgr` data. We first test the effect of the interaction.

```{r}
mgr <- mgr %>%
  unite(species_genotype, species, genotype, sep = "_", remove = F)
lmer_mgr <- lmer(mgr ~ species_genotype * temperature + (1 | species_genotype / isolate), data = mgr, REML = FALSE)
Anova(lmer_mgr)
```

The interaction was significant. We now create different data sets for each temperature and test the effect of `species_genotype` within each temperature.

```{r}
# 15 C
mgr15 <- mgr %>%
  filter(temperature == "15")

lmer_mgr15 <- lmer(mgr ~ species_genotype + (1 | species_genotype / isolate), data = mgr15, REML = FALSE)

Anova(lmer_mgr15)

# 30 C
mgr30 <- mgr %>%
  filter(temperature == "30")

lmer_mgr30 <- lmer(mgr ~ species_genotype + (1 | species_genotype / isolate), data = mgr30, REML = FALSE)

Anova(lmer_mgr30)
```

We can see that there are no effect of `species_genotype` in the `mgr` within each of the temperatures.

## Sporulation & Germination

In this experiment, the production of macroconidia and further germination of 20 randomly selected spores of the different strains of the species-genotype combination, was evaluated in two runs of the experiment.

### Import

```{r message=FALSE, warning=FALSE}
spor <- read_csv(here("data", "spor_germ.csv"))

# again, we create the species_genotype variable
spor <- spor %>%
  unite(species_genotype, species, genotype, sep = "_", remove = F)
head(spor)
```

### Visualize

The code below will produce plots for total macroconidia production and percent germinated spores.

```{r message=FALSE, warning=FALSE}
spor1 <- spor %>%
  group_by(species, genotype, isolate) %>%
  summarize(mean_spor = mean(spores_ml))
spor2 <- spor1 %>%
  ggplot(aes(genotype, mean_spor, color = species)) +
  geom_jitter(width = 0.1, size = 3) +
  theme_few() +
  scale_color_viridis(discrete = TRUE) +
  theme(legend.position = c(1, 5), axis.text.x = element_text(angle = 0, hjust = 0.5, size = 7), legend.text = element_text(size = 12, face = "italic"), axis.title = element_text(size = 12)) +
  labs(x = "", y = expression(N.~of~macroconia~x~10 ^ {
    3
  }), color = "Species") +
  ylim(0, 5) +
  guides(colour = guide_legend(nrow = 2))
spor2
```

```{r message=FALSE, warning=FALSE}
germ1 <- spor %>%
  group_by(species, genotype, isolate) %>%
  summarize(mean_germ = mean(germp))
germ <- germ1 %>%
  ggplot(aes(genotype, mean_germ, color = species)) +
  geom_jitter(width = 0.1, size = 3) +
  theme_few() +
  scale_color_viridis(discrete = TRUE) +
  theme(legend.position = "none", axis.text.x = element_text(angle = 0, hjust = 0.5, size = 7), legend.text = element_text(size = 12, face = "italic"), axis.title = element_text(size = 12)) +
  labs(x = "", y = "Germination rate (%)", color = "Species") +
  guides(colour = guide_legend(nrow = 2))
```

### Model

Mixed model for testing the effect of genotype.

```{r message=FALSE, warning=FALSE}
lmer_spor <- lmer(spores_ml ~ genotype + (1 | genotype / isolate), data = spor, REML = FALSE)
```

```{r}
plot(lmer_spor, type = c("p", "smooth"), col = "black")
qqmath(lmer_spor, id = 0.05, col = "black")
```

Model fitted well the data. Let's proceed.

```{r}
Anova(lmer_spor)
```

#### Means comparison

```{r}
medias <- emmeans(lmer_spor, ~genotype)
med <- cld(medias, Letters = LETTERS, alpha = .05)
med <- data.frame(med)
med
```

### Figure 2

We will produce a combo figure for the three variables.

```{r message=FALSE, warning=FALSE}
grid1 <- plot_grid(spor2, germ, labels = c("B", "C"), ncol = 2, align = "hv")

grid5 <- plot_grid(mgr1, grid1, labels = c("A"), rel_heights = c(1, 1), ncol = 1, align = "hv")
ggsave(here("figs", "figure2.png"), grid5, dpi = 600, width = 6, height = 6.2)

grid5
```

## Pathogenicity

The pathogenicity was assessed based on the progress of the symptoms from the central (inoculated) spikelet. The data were transformed to percent severity of the spike (from the proportion of diseased spikelets). There is one observation of severity for each day after inoculation (dai) on individual spike.

### Import

```{r message=FALSE, warning=FALSE}
pathogen <- read_csv(here("data", "pathogenicity.csv")) %>%
  filter(isolate != "test") %>%
  select(-c(8:12))

pathogen
```

### Transform

Since we have spikes as replicates, we will calculate the mean and standard deviation of severity for plotting purposes.

```{r}
# preparing cultivar 194

pathogen_194 <- pathogen %>%
  filter(cultivar == "BRS 194") %>%
  group_by(cultivar, dai, genotype, species) %>%
  summarize(
    mean_sev = mean(sev),
    sd_sev = sd(sev)
  )

# preparing Guamirim

pathogen_Gua <- pathogen %>%
  filter(cultivar == "BRS Guamirim") %>%
  # filter(exp == 2) %>%
  group_by(cultivar, dai, genotype, species) %>%
  summarize(
    mean_sev = mean(sev),
    sd_sev = sd(sev)
  )
```

### Visualize temporal progress

```{r message=FALSE, warning=FALSE}

brs194 <- pathogen_194 %>%
  group_by(dai, genotype, species) %>%
  ggplot(aes(dai, mean_sev, color = species, shape = genotype, group = interaction(species, genotype))) +
  geom_point(position = position_dodge(width = 0.9)) +
  geom_line() +
  geom_errorbar(
    aes(min = mean_sev - sd_sev, max = mean_sev + sd_sev),
    width = 0.2, alpha = 0.3, position = position_dodge(width = 0.9)
  ) +
  theme_few() +
  ylim(0, 55) +
  xlim(0, 25) +
  theme(legend.position = "right", legend.text = element_text(size = 9, face = "italic")) +
  scale_color_viridis(discrete = TRUE) +
  labs(
    shape = "Genotype",
    color = "Species",
    y = "FHB severity (%) ", x = "Days after inoculation",
    color = "species"
  )

brsgua <- pathogen_Gua %>%
  group_by(dai, genotype, species) %>%
  ggplot(aes(dai, mean_sev, color = species, shape = genotype, group = interaction(species, genotype))) +
  geom_point(position = position_dodge(width = 0.9)) +
  geom_line() +
  geom_errorbar(
    aes(min = mean_sev - sd_sev, max = mean_sev + sd_sev),
    width = 0.2,
    alpha = 0.3,
    position = position_dodge(width = 0.9)
  ) +
  theme_few() +
  theme(legend.position = "none") +
  scale_color_viridis(discrete = TRUE) +
  ylim(0, 55) +
  xlim(0, 25) +
  labs(
    shape = "Genotype",
    color = "Species",
    y = "FHB severity (%) ",
    x = "Days after inoculation",
    color = "species"
  )

fig3 <- plot_grid(brsgua, brs194, labels = c("A", "B"), rel_widths = c(0.7, 1), ncol = 2, align = "hv")
ggsave(here("figs", "figure3.png"), fig3, dpi = 600, width = 9, height = 4)

fig3
```

### Visualize AUDPC

Let's calculate the area under the curve for each isolate. For this, we will use the `audpc` function of the `agricolae` package. First, we need to group several variables up to `dai` (days after inoculation). To make our life easier, lets use the `do` in conjunction with the `tidy` functions of the `dplyr` and `broom` package, respectively. This will calculate the audpc for each spike.

#### BRS 194

```{r}
audpc_194 <- pathogen %>%
  filter(cultivar == "BRS 194") %>%
  # filter(exp == 2) %>%
  unite(species_genotype, species, genotype, sep = "_", remove = F) %>%
  select(
    exp, cultivar, dai, isolate, species_genotype, species, genotype, spike,
    sev
  ) %>%
  group_by(exp, isolate, species_genotype, species, genotype, spike, dai) %>%
  summarize(mean_sev = mean(sev)) %>%
  do(tidy(audpc(.$mean_sev, .$dai)))
names(audpc_194)[8] <- "audpc"
```

Now we can plot the means and standard deviation for each isolate.

```{r}
audpc_194 %>%
  group_by(isolate, species, genotype) %>%
  summarize(
    mean_audpc = mean(audpc),
    sd_audpc = sd(audpc)
  ) %>%
  ggplot(aes(reorder(isolate, mean_audpc), mean_audpc, color = genotype, shape = species)) +
  coord_flip() +
  theme_few() +
  geom_point(position = position_dodge(width = 0.4)) +
  geom_errorbar(aes(min = mean_audpc - sd_audpc, max = mean_audpc + sd_audpc), position = position_dodge(width = 0.4), width = 0.2)
```

#### BRS Guamirim

```{r}
audpc_gua <- pathogen %>%
  filter(cultivar == "BRS Guamirim") %>%
  # filter(exp == 2) %>%
  unite(species_genotype, species, genotype, sep = "_", remove = F) %>%
  select(
    exp, cultivar, dai, isolate, species_genotype, species, genotype, spike,
    sev
  ) %>%
  group_by(exp, isolate, species_genotype, species, genotype, spike, dai) %>%
  summarize(mean_sev = mean(sev)) %>%
  do(tidy(audpc(.$mean_sev, .$dai)))
names(audpc_gua)[8] <- "audpc"
```

Now we can plot the means and standard deviation for each isolate.

```{r}
audpc_gua %>%
  group_by(isolate, species, genotype) %>%
  summarize(
    mean_audpc = mean(audpc),
    sd_audpc = sd(audpc)
  ) %>%
  ggplot(aes(reorder(isolate, mean_audpc), mean_audpc, color = genotype, shape = species)) +
  coord_flip() +
  theme_few() +
  geom_point(position = position_dodge(width = 0.4)) +
  geom_errorbar(aes(min = mean_audpc - sd_audpc, max = mean_audpc + sd_audpc), position = position_dodge(width = 0.4), width = 0.2)
```

### Mixed model

We fit a separate model for each cultivar because the experiments were conducted at different times. We fit a multilevel to test the effect of `species_genotype` as fixed effects and spikes nested within isolates as random effects. The species_genotype interaction was non-significant (not shown), and then a simpler model is used.

#### BRS Guamirim

```{r}
mix_gua <- lmer(
  audpc ~ genotype +
    (1 | isolate / spike),
  data = audpc_gua, REML = FALSE
)

mix_gua
```

Evaluate the model

```{r message=FALSE, warning=FALSE}
plot(mix_gua, type = c("p", "smooth"), col = "black")
qqmath(mix_gua, id = 0.05, col = "black")
```

The model seems OK. Let's check the significance of the species_genotype and estimated means.

```{r}
Anova(mix_gua)
means <- emmeans(mix_gua, ~ genotype)
cld(means)
```

We can see that the variation among isolates was too large and so no difference could be detected among the five species_genotype in the moderate resistant cultivar.

#### BRS 194

We now apply the same model for the susceptible cultivar.

```{r}
mix_194 <- lmer(
  audpc ~ genotype +
    (1 | isolate / spike),
  data = audpc_194
)
mix_194
```

Evaluate the model.

```{r}
plot(mix_194, type = c("p", "smooth"), col = "black")
qqmath(mix_194, id = 0.05, col = "black")
```

Note: I used both the original and the log-transformed `audpc` and the results were the same. Hence, I kept the original data.

```{r}
Anova(mix_194)
means <- emmeans(mix_194, ~ genotype)
cld(means)
```

Again, the effect of species_genotype was not significant due to large variation among isolates.

## Fungicide sensitivity

The response variable is the EC50 estimated in two experiments (true replicates) for each selected isolate from each species_genotype.

### Import

```{r message=FALSE, warning=FALSE}
ec50 <- read_csv(here("data", "ec50.csv"))
ec50
```

### Summarize

Let's obtain the mean and standard deviation of ec50 for each strain.

```{r message=FALSE, warning=FALSE}
ec502 <- ec50 %>%
  unite(species_genotype, species, genotype, sep = "_", remove = F) %>%
  group_by(isolate, species, genotype, species_genotype) %>%
  summarize(
    mean_ec50 = mean(ec50),
    sd_ec50 = sd(ec50)
  )
ec502
```

### Visualize

Let's have a look at the mean EC50 for each isolate.

```{r message=FALSE, warning=FALSE}
ec502 %>%
  ggplot(aes(reorder(isolate, mean_ec50), mean_ec50, shape = species, color = genotype)) +
  geom_point() +
  geom_errorbar(aes(min = mean_ec50 - sd_ec50, max = mean_ec50 + sd_ec50), width = 0.2) +
  theme_few() +
  coord_flip() +
  theme(
    legend.position = "right", axis.text.y = element_blank(),
    legend.text = element_text(size = 7, face = "italic")
  ) +
  ylim(0, 1.8) +
  scale_color_viridis(discrete = TRUE) +
  labs(y = (expression(paste("EC"[50], " (", mu, "g/ml)"))), x = "FGSC strain", shape = "Species", color = "Genotype") +
  ggsave(here("figs", "figure4.png"), dpi = 600, width = 4.5, height = 3)
```