### 2018-01-21 data.Rmd

% Data

## Raw data

The data for this study were obtained from five different (mostly repeated in time) experiments, and were organized in a single `xlsx` file. Data from each experiment are in a separate worksheet and were exported to a `csv` file. They are already organized in a rectangular (tidy) format, mostly in the long-format (a raw is an observation). For some experiments, they are in the wide format (observations in multiple columns), so some wrangling will be needed, which is shown within the respective section of the data analysis page, after data import.

## Download

> The data are stored permanently at the Open Science Framework servers
https://osf.io/c2mbr/ 
Identifier: DOI 10.17605/OSF.IO/C2MBR

>Citation:
Del Ponte, E. M. (2018, January 21). Fitness of **Fusarium graminearum** species complex strains from Brazilian wheat. http://doi.org/10.17605/OSF.IO/C2MBR